---
title: "EDS221 Day 8 Data Visualization"
author: "Dalila Lara"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(here)
library(tidyverse)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)

library(sf)
```

### Read in lizard data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

Plot total length vs weight
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() #aes(x = , y = can also be placed inside geom_point())
```

Histogram of total lengths for all lizards in data
```{r}
#geom_historgram(bins = 40) width of bars
plot_1 <- ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.3,
                 linetype = "dotted")
```

Example 2 
```{r}
#shape = 22 (squares), 19 (points), 18 (diamonds)
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(
    shape = 22,
    color = "cyan4",
    fill = "yellow",
    size = 4,
    alpha = 0.4
  )
```

Mapping an aesthetic based on variables 
```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             alpha = 0.5) +
  theme_minimal()
```

Facet wrap for species 
Makes individual graph for each species
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  facet_wrap(~common_name, scales = "free") 
```

Reordering things 
Totaling counts by the common name to make summarized table of the species 
```{r}
#total counts by common name
lizard_counts <- lizards |>
  group_by(common_name) |>
  summarize(
    counts = n()
  )

#gives exact output as above 
lizards_counts <- lizards |>
  count(common_name)
```

Make a col. graph that plots val. n()
`geom_col()`
`fct_reorder()`
```{r}
#default in ggplot is plot in alphabetical order 
#can't manually specify order of thing, only if factors then use `fct_reorder`
#horizontal chart 
ggplot(data = lizards_counts, aes(y = fct_reorder(common_name, n),
                                 x = n)) +
  geom_col()
```

Synthesis examples 
```{r}
#create a scatter plot, facetted by species for total length vs weight, color and shape of points should differ by species (common name). Change theme, make each faceted pannel have flexible axes limits(scales should be free)

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
             shape = common_name),
             alpha = 0.5,
             size = 1) +
  facet_wrap(~common_name, scales = "free") +
  theme_minimal() +
  labs(x = "Total length (mm)",
       y = "Weight (grams)",
       title = "Jornada Basin lizard sizes",
       subtitle = "All lizards follow the standard length weight relationship with Western whiptails as largest.",
       caption = "Data Jornada Basin lizards")
```

Scales 
  - Qualitative scales distinguish groups
  - Sequenntial scales indicate values or value order
  - diverging scales when there's an obvious mid point and you want to show how much higher or lower things are from it 
  - highlights point something of interest 
  
### fun with scales 

```{r}
#2 color gradient
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high = "orange")
```

```{r}
#mutlicolor gradient
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue", "green", "yellow"))
```

```{r}
#scale_color_steps
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black")

#break points specified with thresholds
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",
                     mid = "white",
                     high = "orange",
                     midpoint = 150,
                     breaks = c(50, 75, 150))
```

```{r}
#breaks specified
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, 
                                  to = 300, 
                                  by = 60))

```

Horizontal Box plot 
Wrangling and ggplot 
```{r}
#y common name and x total length fill color changes on common name 
#factor reorder outside of ggplot 
#data wrangling outside ggplot

#put in order according to median length by species common name
lizards_median <- lizards |>
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median ))

#from lowest median to highest median
levels(lizards_median$common_name)      

#ordered and removes legend
ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE)

#not ordered
ggplot(data = lizards, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name))

#color palettes 
#ordered and removes legend and scale_fill_palettteer_d()
ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "beyonce::X77") +
  theme_minimal()

#view(palettes_d_names)
```

```{r}
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  theme_minimal() +
  #red grid line
  #theme(panel.grid = element_line(color = "red"))

  #only x grid lines are red
  theme(panel.grid.major.x = element_line(color = "red"),
  #y minor grid green color
  #panel.grid.minor.y = element_line(color = "green"))
        panel.grid.minor.y = element_blank(), 
  #change axis title colors
        axis.title.x = element_text(color = "purple", size = 10),
        axis.title.y = element_text(color = "green", size = 10),
  #change text size
        text = element_text(size = 16),
  #plot background
  #adding line to graph
        panel.background = element_rect(color = "red", fill = "cyan3")) +
        annotate("text", x = 300, y = 50, label = "Wooo", color = "blue") +
        geom_vline(xintercept = 250, linetype = "dashed", color = "pink", size = 4)
```

Repulsive labels with `ggrepel`

```{r}
#make subsets with specified site and lizards
wws_lizards <- lizards |>
  filter(common_name == "western whiptail",
         site == "sand")
```

```{r}
# `geom_text_repel`
ggplot(data = wws_lizards, 
       aes(x = total_length,
           y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num)) 
```

```{r}
#view(gapminder)
gapminder |>
  filter(year == 2002,
         continent == "Europe") |>
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```

`gghighlight`
```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) +
  geom_point() +
  gghighlight(weight > 30, label_key = toe_num)
```

### MAPS Make a map
```{r}
# read in sf data 
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot( data = jornada_veg) +
  geom_sf(aes(fill = Name),
          color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```















